---
title: "ggplot2"
author: "Martha Mulugeta"
date: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

devtools::install_github("thomasp85/patchwork")

library(tidyverse)
library(ggridges)
library(patchwork)
```

##create the weather data, cache saves output of code chunk in directory and doesnt re-download dataset each time; dont worry about it
```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## making new plots
start with an old plot! 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

add lables: 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  )
```

x axis tick marks etc 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5", "20")
  ) +
  scale_y_sqrt()
```

##colors

change default colors that ggplot uses; use sclae color function; don't really need to do this
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  ) +
  scale_color_hue(
    name = "Weather Station",
    h = c(50, 250)
  )
```

viridis package
```{r}
ggp_base = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  ) +
   viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base
```

## Overarching themes go first (i.e. bw, minimal), then tweaks in theme go next 

```{r}
ggp_base +
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r}

```




