---
title: "ggplot2"
author: "Martha Mulugeta"
date: "10/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

devtools::install_github("thomasp85/patchwork")

library(tidyverse)
library(ggridges)
library(patchwork)
```

##create the weather data, cache saves output of code chunk in directory and doesnt re-download dataset each time; dont worry about it
```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## making new plots
start with an old plot! 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

add lables: 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  )
```

x axis tick marks etc 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5", "20")
  ) +
  scale_y_sqrt()
```

##colors

change default colors that ggplot uses; use sclae color function; don't really need to do this
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  ) +
  scale_color_hue(
    name = "Weather Station",
    h = c(50, 250)
  )
```

viridis package
```{r}
ggp_base = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    caption = "Data from the rnoaa package"
  ) +
   viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base
```

## Overarching themes go first (i.e. bw, minimal), then tweaks in theme go next 

```{r}
ggp_base +
  theme_minimal() +
  theme(legend.position = "bottom")

```

## more than one dataset 
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp)) +
  geom_line(data = central_park)
  
```

```{r}
waikiki %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(alpha = 0.5, color = "blue")
```

##multi-panel plots
```{r}
ggp_scatter = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_density = 
  weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_density()

ggp_box =
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax) +
  geom_boxplot()

ggp_scatter + ggp_density

ggp_scatter + (ggp_density + ggp_box)
```

##Data manipulation
```{r}
weather_df %>%
  mutate(
    name = factor(name),
    name = fct_relevel(name, "Waikiki_HA", "CentralPark_NY")
  ) %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_boxplot()
```

reorder instead of relevel 
```{r}
weather_df %>%
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_boxplot()
```

##restructure then plot
```{r}
weather_df %>%
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = observation)) + 
  geom_density(alpha = 0.5) +
  facet_grid(~name) +
  theme(legend.position = "bottom")
```

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 
```

```{r}
fas_data %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) %>% 
  drop_na(dose, day_of_tx, pn_day) %>% 
  mutate(
    outcome = factor(outcome),
    outcome = fct_reorder(outcome, pn_day)
  ) %>% 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
facet_grid(day_of_tx ~ outcome)
```






