---
title: "ggplot 1"
author: "Martha Mulugeta"
date: "9/26/2019"
output: github_document
---

This is the repo for examples related to Viz and EDA
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```

##create the weather data
```{r load_data, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

##create a ggplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

alternate way of making this plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

saving initial plots. mostly don't do this 
```{r}
weather_df %>% filter(name == "CentralPark_NY")
```

```{r}
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

scatterplot
```

adding color (corresponding to the names; alpha adds transparency to the points on a plot)
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.4)
```

why do 'aes' positions matter? geom_plot is creating curve going through middle of all data points; exploratory
se = FALSE turns standard errors of curve off

first
first aes applies to everything coming after; color matters inside geom point not geom smooth
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.4) +
  geom_smooth(se = FALSE)
```
vs

here, color for individual plots and smooth curves for individuals 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE)
```

facets!
want names to be in the columns; separate plots for each name
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

this is fine, but not interesting.
size = prcp means that the greater the precipitation, the larger the point
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(size = 2, se = FALSE)
```

##some extra stuff
just the curves; inferior to having the points
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(size = 2, se = FALSE)
```

2d density
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_bin2d()

# install.pachages ('hexbin') to use geom_hex
```

##more kinds of plots
position = dodge makes the bars next to each other instead of stacked 

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(~name)
```

density plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3)
```

univariate plots: boxplots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

univariate plots: violin plots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

ridge plots!!!
Unstacked version of geom_density 
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```
connect each point in the ggplot with a '+' and within manipulation with ' %>%'

##saving a plot
```{r}
ggp_ridge_temp =
  weather_df %>% 
    ggplot(aes(x = tmax, y = name)) +
    geom_density_ridges()

ggsave("ggplot_ridge_temp.pdf", ggp_ridge_temp)
```

embedding plots in rmarkdown
```{r, fig.width = 6}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE) 
```

```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE) 
```




